Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-06-18 doc/ms.tex
Include udpated captions for spectroscopy figures.

% \author{A.~A.~Miller}
% \affil{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA) and Department of Physics and Astronomy, Northwestern University, 2145 Sheridan Road, Evanston, IL 60208, USA}
% \affil{The Adler Planetarium, Chicago, IL 60605, USA}
%
% \author[0000-0002-8036-8491]{Yi Cao}
% \affil{eScience Institute and Astronomy Department, University of Washington,
%   Seattle, WA 98195}
%
% \author{B.~D.~Bue}
% \affil{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}
%
% \author{M.~M.~Kasliwal}
% \affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{R.~Ferretti}
% \affil{The Oskar Klein Centre, Department of Physics, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{A.~Goobar}
% \affil{The Oskar Klein Centre, Department of Physics, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{R.~R.~Laher}
% \affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{F.~J.~Masci}
% \affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{P.~E.~Nugent}
% \affil{Lawrence Berkeley National Laboratory, Berkeley, California 94720, USA}
% \affil{University of California -- Berkeley, Berkeley, CA 94720, USA}
%
% \author{J.~Sollerman}
% \affil{The Oskar Klein Centre, Department of Astronomy, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{F.~Taddia}
% \affil{The Oskar Klein Centre, Department of Astronomy, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{S. R. Kulkarni}
% \affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
%
\author{The Alphabet Friends}
\affil{the intermediate Palomar Transient Factory}

@@

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a 
stars with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass \citeauthor{2012ApJ...744L..17B} derive the mean density of the 
progenitor star, confirming that at least some type Ia SNe come from 
compact and degenerate stars.

@@

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.45\textwidth]{Chi2.pdf}
  \includegraphics[width=0.45\textwidth]{VelocityPlot.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}.
    \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The red
    vertical line and the orange region indicate $t_0$ and its
    3-$\sigma$ confidence interval from Section
    \ref{sec:lc_fit}.
    \textit{Right panel:} Observed \ion{Si}{2}\,6355
    velocities and the best-fit power-law velocity with an index of
    $-0.22$.}
  \label{fig:velocity_t_exp}
\end{figure*}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{CarbonFeature.pdf}
  \caption{Evolution of the \ion{C}{2} features observed in the 
  early spectra of \abc. The raw spectra are shown in grey, while 
  the solid black lines show a smoothed version to highlight the 
  observed spectral features.}
  \label{fig:carbon}
\end{figure}

@@

Qualitatively, the PDD models provide a better match to the observations of \abc\ than DD models. In particular, PDD models provide higher luminosities and faster rise times in the days after explosion, a more rapid evolution towards blue optical colors, and the formation of strong \ion{C}{2} lines that gradually disappear in the $\sim$week after explosion. Quantitatively, there are still some short comings of the models presented in \citet{2014MNRAS.441..532D}. In particular, the power-law index for the $g$-band rise is $\alpha \approx 3$ for each of the PDD models, which is significantly more steep than $\alpha = 0.96$, which we measure for \abc. The optical colors for \abc\ are blue from our first epoch $\sim$1.7 d after explosion and remain approximately constant in the 10 d after explosion, whereas the PDD models in \citet{2014MNRAS.441..532D} either exhibit distinct color evolution (i.e.\ are not constant) or are too red in the first few days after explosion. Nevertheless, the PDD models presented in \citet{2014MNRAS.441..532D} provide several attractive explanations for the unusual features in the early behavior of \abc, and it may be possible that small adjustments to the model, e.g., additional mixing of the radioactive products from the explosion, can better match \abc.

@@

\amiller{Need to thank Rollin for chatting about \ion{C}{2}}
\amiller{Need to thank Uli, Rahman for chatting about SALT2}

% Observers: Michael West, Gemini observers, J. Cohen, Plant,J.Torres,Jang,Cunningham,Guhathakurta, Ravi, Cao, Yan, Vedantham, Kulkarni, N. Suzuki, Y. Cao, V. Ravi, Walters, Lunnan, Ho, De

YC thanks supports from the postdoc fellowship in the eScience
institute, University of Washington.

Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA.

DAH, CM, and GH are supported by NSF-1313484. Support for IA was provided by NASA through the Einstein Fellowship Program, grant PF6-170148. This work makes use of observations from the LCO network.

>>>>

% \author{A.~A.~Miller}
% \affil{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA) and Department of Physics and Astronomy, Northwestern University, 2145 Sheridan Road, Evanston, IL 60208, USA}
% \affil{The Adler Planetarium, Chicago, IL 60605, USA}
%
% \author[0000-0002-8036-8491]{Yi Cao}
% \affil{eScience Institute and Astronomy Department, University of Washington,
%   Seattle, WA 98195}
% \affil{Google, Mountain View, CA 94043, USA}
%
% \author{B.~D.~Bue}
% \affil{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}

% \author{R.~Ferretti}
% \affil{The Oskar Klein Centre, Department of Physics, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{A.~Goobar}
% \affil{The Oskar Klein Centre, Department of Physics, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{G.~Hosseinzadeh}
% \affil{Las Cumbres Observatory, Goleta, CA 93117, USA}
% \affil{Physics Department, University of California, Santa Barbara, CA 93106, USA}
%
% \author{M.~M.~Kasliwal}
% \affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{R.~R.~Laher}
% \affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{F.~J.~Masci}
% \affil{Infrared Processing and Analysis Center, California Institute of Technology, Pasadena, CA 91125, USA}
%
% \author{P.~E.~Nugent}
% \affil{Lawrence Berkeley National Laboratory, Berkeley, California 94720, USA}
% \affil{University of California -- Berkeley, Berkeley, CA 94720, USA}
%
% \author{J.~Sollerman}
% \affil{The Oskar Klein Centre, Department of Astronomy, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{F.~Taddia}
% \affil{The Oskar Klein Centre, Department of Astronomy, AlbaNova, SE-106 91 Stockholm, Sweden}
%
% \author{S. R. Kulkarni}
% \affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
%
\author{The Alphabet Friends}
\affil{the intermediate Palomar Transient Factory}

@@

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a 
star with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass \citeauthor{2012ApJ...744L..17B} derive the mean density of the 
progenitor star, confirming that at least some type Ia SNe come from 
compact and degenerate stars.

@@

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.45\textwidth]{Chi2.pdf}
  \includegraphics[width=0.45\textwidth]{VelocityPlot.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}.
    \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The blue
    vertical line and the orange shaded region indicate $t_0$ and its
    99.7\% confidence interval from Section
    \ref{sec:lc_fit}, respectively.
    \textit{Right panel:} Observed \ion{Si}{2}\,$\lambda$6355
    velocities (blue circles) and the best-fit power-law model
     with an index of $-0.22$ (dashed line). Additionally the 
     measured velocites of \ion{C}{2}\,$\lambda\lambda$6580, 7234 
     are shown.}
  \label{fig:velocity_t_exp}
\end{figure*}

@@

\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{CarbonFeature.pdf}
  \caption{Evolution of the \ion{C}{2} features observed in the 
  early spectra of \abc. The raw spectra are shown in orange, while 
  the solid blue lines show the best-fit models including gaussian 
  components for each line and a linear component for the continuum 
  (see text for further details). The dark grey vertical lines show 
  the measured line centers, indicating the decline in the 
  photosphere velocity in the $\sim$wk after explosion (note that 
  \ion{C}{2}\,$\lambda$7234 is not detected in the $-10.9$ d 
  spectrum). The phase of each spectrum relative to 
  $T_{B,\mathrm{max}}$ is shown to the right of each spectrum.
  }
  \label{fig:carbon}
\end{figure}

@@

Qualitatively, the PDD models provide a better match to the observations of \abc\ than DD models. In particular, PDD models provide higher luminosities and faster rise times in the days after explosion, a more rapid evolution towards blue optical colors, and the formation of strong \ion{C}{2} lines that gradually disappear in the $\sim$week after explosion. Quantitatively, there are still some short comings of the models presented in \citet{2014MNRAS.441..532D}. In particular, the power-law index for the $g$-band rise is $\alpha \approx 3$ for each of the PDD models, which is significantly more steep than $\alpha = 0.98$, which we measure for \abc. The optical colors for \abc\ are blue from our first epoch $\sim$1.7 d after explosion and remain approximately constant in the 10 d after explosion, whereas the PDD models in \citet{2014MNRAS.441..532D} either exhibit distinct color evolution (i.e.\ are not constant) or are too red in the first few days after explosion. Nevertheless, the PDD models presented in \citet{2014MNRAS.441..532D} provide several attractive explanations for the unusual features in the early behavior of \abc, and it may be possible that small adjustments to the model, e.g., additional mixing of the radioactive products from the explosion, can better match \abc.

@@

We are deeply grateful to R.~C.~Thomas for induldging a slew of questions regarding \ion{C}{2} in SNe. Similarly, it is our pleasure to buy a beer for R.~Amanullah and U.~Feindt for discussions regarding SALT2.

Much of the analysis presented herein would not have been possible without the help of several observers. We thank M.~West for taking the first spectrum of \abc\ as a ToO on the DCT. We also thank the Gemini service observers for executing our ToO observations. Additionally, J.~Cohen, N.~Suzuki, V.~Ravi, R.~Walters, A.~Ho, and L.~Yan helped obtain data for this paper.

% Observers:  Plant,J.Torres,Jang,Cunningham,Guhathakurta, Vedantham, De

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of the Data Science Fellowship Program. YC thanks support from a postdoc fellowship in the eScience institute, University of Washington.

DAH, CM, and GH are supported by NSF-1313484. Support for IA was provided by NASA through the Einstein Fellowship Program, grant PF6-170148.

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network.
